\chapter{Generalising the Charney–Phillips staggering for arbitrary meshes}
\label{ch:cp}

The Charney--Phillips vertical staggering of variables \citep{charney-phillips1953} is suitable for structured meshes with cells stacked in columns.
This staggering has been adopted by several operational models \citep{davies2005,girard2014,yang2007} because it avoids the computational mode that is associated with the Lorenz vertical staggering \citep{arakawa-konor1996}.
The generalisation of the Lorenz staggering for arbitrary meshes is straightforward \citep{weller-shahrokhi2014} but this is not true for the Charney--Phillips staggering.
On a finite volume mesh, variables are ordinarily placed at cell centres or cell faces. In the Charney--Phillips staggering, the thermodynamic variable is placed at only those cell faces that lie on the vertical coordinate surfaces, and vertically-oriented faces have no thermodynamic information.
This existing staggering is unsuitable for arbitrary finite volume meshes because faces can have any orientation.

\TODO{somewhere give evidence for Lorenz computational mode having been excited \citep{girard2014,yi-park2017}}

\input{cp/schaerWaves.tex}

\section{Advective-form transport for arbitrary Charney--Phillips meshes}

\TODO{describe the CP advection scheme: interpGrad+fvcReconstructCP (advective form)}

\input{cp/arakawaKonor.tex}

\section{Horizontal transport on distorted Charney--Phillips meshes}
\TODO{
\begin{itemize}
	\item Test: horizontal advection; advectiveFoamF, advectionFoam with linear interp, advectionFoam with linearUpwind interp; measure l2 and linf errors (s.t. they can be compared with cell-centred, C-grid results); uniform and edgeGraded meshes
	\item Plot: error fields, analytic+numerical contours of final solution, which schemes and which meshes? too many combinations to plot all of them
	\item Plot: time-series of conservation, ExnerFoamH and ExnerFoamCP, uniform mesh, horizontal edgeGrading, vertical edgeGrading
	\item Conclusion: advection is conservative on uniform meshes, non-conservative on non-uniform meshes
	\item Conclusion: advectiveFoamF errors are comparable to advectionFoam with linear interpolation?
	\item Conclusion: advectiveFoamF errors are worse than advectionFoam with linearUpwind interpolation?
\end{itemize}
}

