\chapter{Numerically stable transport over steep slopes}
\label{ch:cubicFit}

\begin{highlights}
{\Large Highlights}
\begin{itemize}
	\item The new cubicFit transport scheme is second-order convergent regardless of mesh distortions or the choice of velocity field
	\item Sub-grid reconstructions are mostly precomputed depending on the mesh geometry alone
	\item Misalignment of the velocity field with mesh layers is the primary source of numerical error, not simply mesh distortions
\end{itemize}
\end{highlights}

A huge variety of transport schemes have been developed for atmospheric models, but few are able to account for distortions associated with steep terrain because they treat horizontal and vertical transport separately \citep{kent2014}, resulting in numerical errors called `splitting errors'.
Such errors can be reduced by explicitly accounting for transverse fluxes when combining fluxes \citep{leonard1996}, but splitting errors are still apparent in flows over steep terrain where meshes are highly distorted and metric terms in a terrain-following coordinate transform are large \citep{chen2017}.

\TODO{I haven't stated the order of convergence for any schemes mentioned here}
Transport schemes are often classified as dimensionally-split or multidimensional.
Dimension\-ally-split schemes such as \citep{lin-rood1996,guo2014} calculate transport in each dimension separately before the flux contributions are combined.  Such schemes are computationally efficient and allow existing one-dimensional high-order methods to be used.
When dimensionally-split schemes are used for horizontal transport, quadrilateral meshes are needed because the mesh dimensions are inherently separable.  Special treatment is required at the corners of cubed-sphere panels where local coordinates differ \citep{putman-lin2007,guo2014}.
Dimensional splitting is often used for vertical transport and, for similar reasons, dimensionally-split schemes have only been used with terrain-following coordinate transforms and not cut cells.
Perhaps confusingly, dimensionally-split schemes are sometimes called multidimensional, too, because they use one-dimensional techniques for multidimensional transport.

Unlike dimensionally-split schemes, multidimensional schemes consider transport in two or three dimensions together.
There are several subclasses of multidimensional schemes that include
semi-Lagrangian finite volume schemes (also called conservative mesh remapping),
swept-area schemes (also called flux-form semi-Lagrangian, incremental remapping, or forward-in-time),
and method-of-lines schemes (also called Eulerian schemes).
Two-dimensional semi-Lagrangian finite volume schemes such as \citep{iske-kaeser2004,lauritzen2010} integrate over departure cells that are found by tracing backward the trajectories of cell vertices.  These schemes are conservative because departure cells are constructed so that there are no overlaps or gaps, which requires that cell areas are simply-connected domains \citep{lauritzen2011book}.
SLICE-3D is a three-dimensional semi-Lagrangian finite volume scheme for latitude-longitude meshes that applies separate conservative remappings in each dimension \citep{zerroukat-allen2012}.
Swept-area schemes such as \citep{lashley2002,skamarock-menchaca2010,lauritzen2011,thuburn2014} calculate the flux through a cell face by integrating over the upstream area that is swept out over one time-step.  Such schemes differ in their choice of area approximation, sub-grid reconstruction, and spatial integration method.
Because swept-area schemes integrate over the reconstructed field, they typically require a matrix-vector multiply per face per time-stage \citep{thuburn2014,skamarock-menchaca2010}.
Method-of-lines schemes such as \citep{weller2009,skamarock-gassmann2011} use a spatial discretisation to reduce the transport PDE to an ODE that is typically solved using a multi-stage time-stepping method.
A method-of-lines scheme using a spectral element reconstruction was recently developed to achieve accurate solutions near the surface of cut cell meshes \citep{steppeler-klemp2017}.
Unlike semi-Lagrangian finite volume schemes, swept-area and method-of-lines schemes achieve conservation for small-scale rotational flows.
Such flows can twist the departure domain to such an extent that the domain intersects itself \citep{lauritzen2011}.  In two dimensions, a self-intersecting departure domain has a bowtie or hourglass shape.
There are many more types of atmospheric transport schemes, but all can be classified according to their treatment of the three spatial dimensions.  A more comprehensive overview is presented by \citet{lauritzen2014}.

For transport schemes that are ordinarily classified as `multidimensional', a further distinction ought to made between horizontally-multidimensional and three-dimensional schemes.
Most multidimensional schemes are only horizontally-multidimensional because, while the two horizontal dimensions are considered together, horizontal and vertical transport are still treated separately.
This separate treatment becomes less justifiable as atmospheric models are using increasingly fine horizontal mesh spacings that resolve small-scale steep slopes, resulting in greater mesh distortion and possible splitting errors \citep{kent2014}.
Three-dimensional schemes avoid any splitting errors over steep slopes, but only a few conservative three-dimensional schemes have been used in atmospheric models.
The multi-moment constrained finite volume scheme \citep{ii-xiao2009} is a three-dimensional scheme that has been used to simulate nonhydrostatic flows over orography with terrain-following coordinates on a $x$--$z$ plane \citep{li2013}.  Simulations of subcritical flow around a cylinder have also been performed on a three-dimensional hexahedral-prismatic hybrid mesh \citep{xie-xiao2016}.
The Multidimensional Positive Definite Advection Transport Algorithm (MPDATA) is another three-dimensional scheme that is suitable for arbitrary meshes.
\TODO{what about its computational characteristics?}
It has been used on triangular unstructured meshes to simulate two-dimensional nonhydrostatic flows over orography \citep{smolarkiewicz-szmelter2011}, and in three-dimensional transport tests \citep{smolarkiewicz-szmelter2005}. 
Most recently, \citet{kuehnlein-smolarkiewicz2017} extended MPDATA to enable semi-implicit integrations of the compressible Euler equations on arbitrary meshes.
The three-dimensional method-of-lines scheme developed by \citet{weller-shahrokhi2014} has been used in two-dimensional flows over orography on Cartesian $x$--$z$ planes with distorted meshes \citep{chen2017}.
This finite volume scheme uses a moving weighted least-squares reconstruction \citep{lashley2002,thuburn2014} that makes it suitable for arbitrary meshes.
Similar least-squares approaches have been applied previously to shallow water flows \citep{cuetofelgueroso2006}, aeronautic \citep{cuetofelgueroso2007} and porous media \citep{white2017} simulations.

This chapter presents a new multidimensional method-of-lines scheme, `cubicFit', that improves the stability of the scheme by \citet{weller-shahrokhi2014} and avoids all splitting errors.  To reconstruct values at cell faces, the scheme fits a multidimensional cubic polynomial over an upwind-biased stencil using a least-squares approach.
The implementation uses stability conditions derived from a von Neumann stability analysis to select appropriate polynomial fits for stencils in highly-distorted mesh regions.
This stabilisation procedure has similarities to the Multidimensional Optimal Order Detection (MOOD) method \citep{clain2011,diot2013}.
However, MOOD is an \textit{a posteriori} method that detects discontinuities in the solution.  In contrast, the cubicFit stabilisation procedure is an \textit{a priori} method that depends upon the mesh geometry only, and reconstruction weights can be precomputed without knowledge of the velocity field or tracer field.
\TODO{another alternative might be matrix regularisation... a regularised moving least squares technique suitable for poor-quality meshes is formulated by \citet{wang2018}, but hasn't been applied to a PDE, just used for surface reconstruction?}
Hence, the computational cost of the cubicFit scheme is lower than most swept-area schemes that require a matrix-vector multiply per face per time-stage \TODO{how big is the matrix?}.  Instead, the computational cost is more comparable to dimensionally-split schemes, with the cubicFit scheme requiring only $n$ multiplies per face per time-stage where $n$ is the size of the stencil.

The remainder of this chapter is organised as follows.
Section~\ref{sec:cubicFit:transport} starts by discretising the transport equation using a method-of-lines approach before describing the cubicFit transport scheme, and a multidimensional linear upwind transport scheme that is included in the OpenFOAM software distribution \citep{openfoam-numerics}.
Subsequent sections evaluate the cubicFit scheme in a series of three idealised numerical tests.
The test in section~\ref{sec:cubicFit:schaerAdvect} follows \citet{schaer2002}, transporting a tracer horizontally above steep mountains on two-dimensional, highly-distorted terrain-following meshes.
Section~\ref{sec:cubicFit:tfAdvect} formulates a new tracer transport test that uses a terrain-following velocity field to challenge transport schemes on cut cell meshes.
Finally, in section~\ref{sec:cubicFit:deformationSphere}, we assess the cubicFit transport scheme on hexagonal-icosahedral meshes and cubed-sphere meshes using a standard test of deformational flow on a single-layer spherical Earth, as specified by \citet{lauritzen2012}.

\input{cubicFit/transport}
\input{cubicFit/scheme}
\input{cubicFit/linearUpwind}
\input{cubicFit/schaerAdvect.tex}
\input{cubicFit/tfAdvect.tex}
\input{cubicFit/deformationSphere.tex}

